---
title: "EDS 213 - Assignment 2.1 and 2.2"
author: "Steven Cognac"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dataRetrieval) 
library(tidyverse)
library(metajam)
library(here)
library(patchwork)
library(gt)

```

USGS API sources
 - https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html#daily-data
 - https://dashboard.waterdata.usgs.gov/app/nwd/?region=lower48&aoi=default
 
Metajam sources
 - https://brunj7.github.io/EDS-213-metadata/slides/metajam_intro.html#3
 

# **Assignment 2.1 - dataRetrieval**


## 2.1.1 Add the Santa Paula Creek guage to exercise we did in class
## Retrieve USGS stream guage data at the Ventura River and Santa Paula Creek from 2019-10-01 to 2020-10-10
```{r}

siteNumber <- c(11118500, 11113500) # Ventura river and Santa Paula Creek USGS stations
parameterCd <- c("00060") # discharge & gauge height
startDate <- "2021-10-04" # use "" to grab entire data from beginning of recording
endDate <- "2021-10-04"
tz = "America/Los_Angeles"


# create database based on above parameters
discharge <- readNWISuv(siteNumber, parameterCd, startDate, endDate, tz) %>% 
  mutate(river = case_when(site_no == "11118500" ~ "ventura_river",
                           site_no == "11113500" ~ "santa_paula_creek"))

```

## 2.1.2 What is the best way to query multiple sites using this R package?

**Best way to query multiple sites is to put all site numbers in a vector using the concatenate c() function. **

## 2.1.3 Create a plot showing any stream flow response to the rain of last night (10/04). Determine at what time did the stream flow peaked (if any) at the two locations?

 - **Based on the data and graphs, the Ventura River had a slight peak on October 7.  However, the river has a wide floodplain and is dammed less than 1 mile upstream from the Lake Casitas dam. As such, the flow is heavily regulated and minor discharge deviations in the Ventura River are unlikely a result of the rainstorm.**
 - **The Santa Paula Creek is a smaller creek that's not dammed and more likely  influenced by storm events.  Based on the data and graphs, Santa Paula Creek discharge peaked on October 8, 2021. **

below the Santa Paula Creek discharge peaked October 8, four days after the storm event Can't fully answer question **

```{r}

# plot of storm event stream discharge data after October 1, 2021
storm_plot <- ggplot(data = discharge, aes(x = dateTime, y = X_00060_00000, color = river)) +
  geom_line() +
  labs(x = "Date",
       y = "Discharge (cfs)",
       title = "October 4, 2021 rain storm event",
       color = "Legend") +
  geom_vline(xintercept = as.numeric(as.Date("2021-10-04")), color = "red", lwd = 2)


storm_plot
```

# **Assignment 2.2 - metajam **

## 2.2.1 Finish the exercise we started during the lab session to use metajam to download the Alaskan household languages data from DataONE.

```{r, error=TRUE}
# downloading data
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"


# download csv files with metajam to folder titled "metajam" with the here() function 
data_path <- download_d1_data(data_url, here("metajam"))
```

## read in .csv data with metajam
```{r}

# download the data based on the file path with metajam. 
# creates a list of all data, which has pre-defined names
data <- read_d1_files(data_path)

# view metadata
data$attribute_metadata

```


## Write a piece of code that will compute the percentage of Alaskan household speaking only English for the year 2009 through 2015
```{r}
# get the dataframe
hh_data <- data$data
hh_data

# filter data for dates from 2009 to 2015. Group by year and region based on yearly means
alaska_data <- hh_data %>% 
  filter(Year %in% c("2009", "2010", "2011", "2012", "2013", "2014", "2015")) %>% 
  dplyr::rename(Region = "SASAP.Region") %>% 
  group_by(Year) %>% 
  summarise(percent = (mean(speak_only_english, na.rm = TRUE) / (mean(total, na.rm = TRUE))) * 100)
alaska_data

```
# plot of Alaskan household speaking only English for the years 2009 through 2015

```{r}

alaska_standardized <- ggplot(data = alaska_data, aes(x = Year, y = percent)) +
  geom_line() +
  labs(y = "Percent",
       title = "Percentage of English Speaking Households in Alaska") +
  scale_y_reverse()

alaska_standardized

```

